---
title: "Practical Machine Learning course project"
author: "Dan Huang"
date: "27 Apr 2020"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---


1. Data input and R package activation 

```{r, results='hide'}

pml.training <- read.csv("C:/Morphosys/Dan markdown/pml-training.csv", row.names=1)
pml.testing <- read.csv("C:/Morphosys/Dan markdown/pml-testing.csv", row.names=1)

.libPaths("C:/R/R-3.6.1/library")
library("caret")
library("randomForest")

```


2. Data Clearning

```{r}

Train <- pml.training[,colSums(is.na(pml.training))==0] ## to remove columns with NA values 
Test <- pml.testing[, colSums(is.na(pml.testing))==0] ## to remove columns with NA values 
summary(Train) 

```
The majority (> 99%) of observations of variables with prefix of "kurtosis", "skewness", " amplitude", "max" and "min" are missing values, therefore these variables will be removed in prediction.
In addition, variables "raw_timestamp_part_1" and "raw_timestamp_part_2" are not needed for prediction and therefore be removed
```{r}

Train <- Train[, -c(11:19, 42:47, 51:59, 73:81)] 
Train <- Train[, -c(2,3)]
Test <- Test[, -c(2,3)]

```


3. Pre processing

1) To check the nonZeroVar
```{r}

nsv <- nearZeroVar(Train, saveMetrics=TRUE, names = TRUE)
print(nsv)

```
Variable "new_window" will be removed as it is a near Zero Variance predictor


```{r}

Train <- Train[, -c(3)]
Test <- Test[, -c(3)]

```


2) To center the scale of continuous variables

```{r}

preObj <- preProcess(Train[, -c(56)], method = c("center", "scale"))
Train_ <- predict(preObj, Train[, -c(56)])
Trainingset <- cbind.data.frame(Train[, c(56)], Train_)
names(Trainingset)[1] <- "Class"

preObj2 <- preProcess(Test[, -c(56)], method = c("center", "scale"))
Test_ <- predict(preObj2, Test[, -c(56)])
Testingset <- cbind.data.frame(Test[, c(56)], Test_)
names(Testingset)[1] <- "Class"

```


4. Data slicing and Predicting with random forest model 

```{r}

set.seed(2020)

inTrain <- createDataPartition(y=Trainingset$Class, p=0.7, list = FALSE)
Training <- Trainingset[inTrain,]
Validation <- Trainingset[-inTrain,]

modFit <- train(Class~ ., method = "rf",  data=Training, trControl = trainControl(method = "cv", number = 4), importance = TRUE)

```



5. Cross validation and predict for Testing data
```{r}

pred <- predict(modFit, Validation)
print(confusionMatrix(Validation$Class,pred), digits = 3)

```
From the result we can see this model has good performance with >99% accuracy. Therefore we can use it for predicting the testing data.
```{r}

pred <- predict(modFit, Testingset)
pred

```
